---
title: "Computing Assessment"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyr)
library(readr)
library(dplyr)
library(ggplot2)
library(stringr)
library(lubridate)
```

# R Programming Part 1

For this portion of the assessment, we will be working with the MPG dataset. To answer the questions, you will need to complete the following tasks:
- Read the "Auto_MPG_Exam.csv" data. Name your data df_auto. Ensure stringAsFactors=FALSE
- Create a column named "Car_Make" from the first element of the "Car_Name" variable
- Update "horsepower" to be type numeric
- Change the type of "Cylinders", "Model_Year", and "Origin" from numeric to character (Hint: you will need to use the as.character function)

```{r}
df_auto <- read_csv('Auto_MPG_Exam.csv')
```

```{r}
df_auto$Car_Make <- stringr::word(df_auto$Car_Name, 1)
```

```{r}
df_auto$Horsepower <- as.numeric(df_auto$Horsepower)
df_auto$Cylinders <- as.character(df_auto$Cylinders)
df_auto$Model_Year <- as.character(df_auto$Model_Year)
df_auto$Origin <- as.character(df_auto$Origin)
```

## Question 16

How many unique values of "Car_Make" are there? 

```{r}
n_distinct(df_auto$Car_Make)
```

## Question 17

How many variables in "df_auto" are of type character?

```{r}
str(df_auto)
```

We are going to create a function that provides univariate analysis on a provided variable vector, named “data”. Create a function "univariate_analysis" that:
- Calculates the number of observations, mean, standard deviation, and quartiles of a provided vector.
- Ensure the na.rm argument is TRUE when calculating the mean, standard deviation, and the quartiles.
- Have the function return the results in a list with named elements "obs", "mean", "stdev", and "quartiles"

```{r}
univariate_analysis <- function(data) {
  obs = length(data)
  mean = mean(data, na.rm = TRUE)
  stdev = sd(data, na.rm = TRUE)
  quartiles = quantile(data, probs=c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)
  return(list(obs, mean, stdev, quartiles))
}
```


## Question 18

Using the function created, what is the 2nd quartile (the median, the 50th percentile) of "MPG"?

```{r}
univariate_analysis(df_auto$MPG)
```

To the "univariate_analysis" function, create a histogram plot named "histogram" using the code provided below. Add the plot to the return list with the element named "histogram":

histogram <- ggplot(data=data.frame(x=data), aes(x=x)) +  geom_histogram()

```{r}
univariate_analysis <- function(data) {
  obs = length(data)
  mean = mean(data, na.rm = TRUE)
  stdev = sd(data, na.rm = TRUE)
  quartiles = quantile(data, probs=c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)
  histogram <- ggplot(data=data.frame(x=data), aes(x=x)) +  geom_histogram()
  return(list(obs, mean, stdev, quartiles, histogram))
}
```

## Question 19

What is the default number of bins in the histogram for "MPG"?

```{r}
univariate_analysis(df_auto$MPG)
```

Thus far, we have assumed the variable provided will be numeric. However, we should account for categorical variables. Update the "univariate_analysis" function:
- If the provided vector is numeric, execute everything created in the function thus far (Hint: you can use is.numeric(data)==TRUE to evaluate this condition)
- If the provided vector is character (is.character(data)==TRUE):
- Store the number of elements in the vector in a variable named "obs"
- Store the unique values of the vector in a variable named "unique_levels". You can use unique(data) to calculate this.
- Create a bar chart named "bar_chart" using the code provided below. This code plots the number of observations by the unique levels of the variable:
bar_chart <- ggplot(data=data.frame(x=data), aes(x=x)) + geom_bar()
- Returns the previous variables in a list with the elements named "obs", "unique_levels", and "bar_chart")

```{r}
univariate_analysis <- function(data) {
  if(is.numeric(data) == TRUE) {
    obs = length(data)
    mean = mean(data, na.rm = TRUE)
    stdev = sd(data, na.rm = TRUE)
    quartiles = quantile(data, probs=c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)
    histogram <- ggplot(data=data.frame(x=data), aes(x=x)) +  geom_histogram()
    return(list(obs, mean, stdev, quartiles, histogram))
  }
  
  else if(is.character(data) == TRUE) {
    obs = length(data)
    unique_levels = unique(data)
    bar_chart <- ggplot(data=data.frame(x=data), aes(x=x)) + geom_bar()
    return(list(obs, unique_levels, bar_chart))
  }
}
```

## Question 20

Using the function created, what is the median of "Displacement"? (Round your answer to 1 decimal place, example: nn.n)

```{r}
univariate_analysis(df_auto$Displacement)
```

## Question 21

Using the function created, what is the most common number of cylinders for cars in the data set?  

```{r}
univariate_analysis(df_auto$Cylinders)
```

# R Programming Part 2

For this portion of the assessment, we will be working with “traffic_stops_R.csv”. Read the “Traffic_stops_R.csv” data into R.

We want to create a new dataframe that meets the following criteria:
- The new dataframe contains only the violation and weekday column when the stop occurred for female drivers of age greater than 60 that were stopped on a Sunday. To achieve this:
- First create a subset dataframe where you will need to filter the data for female drivers of age greater than 60.
- Next, using your subset dataframe and the library lubridate, add a new column called weekday_of_stop containing the number of the weekday when the stop
occurred. Use the wday(mdy(stop_date)) function to create this new column (Sunday = 1).
- Don’t forget to filter for those drivers that were stopped on a Sunday.

```{r}
traffic_raw <- read_csv('traffic_stops_R.csv')
```

```{r}
traffic <- traffic_raw %>% filter(driver_gender == 'female' & driver_age >60)
dim(traffic)
```

```{r}
traffic <- traffic %>% mutate(weekday_of_stop = wday(mdy(stop_date)))
traffic <- traffic %>% filter(weekday_of_stop != 1)
```

## Question 22

How many observations does your new dataframe contains?

```{r}
dim(traffic)
```

## Question 23

What is the most frequent driving violation in your new dataframe?

```{r}
traffic %>% count(violation)
```



