---
title: "Statistics Lab 9 - Diagnostics Part 2"
output: html_notebook
---

```{r}
library(tidyr)
library(readr)
library(dplyr)
library(stringr)
library(ggplot2)
```

# Question Set 1

For the next few questions, we will revisit the mtcars data set. Use the whole data set (do not subset).

```{r}
mtcars <- read_csv("https://raw.githubusercontent.com/IAA-Faculty/statistical_foundations/refs/heads/master/mtcars.csv")
```

## Question 1

Create the Pearson's correlation between all the quantitative x-variables.  What is the largest value in magnitude of the correlation among the x-variables?

```{r}
mtcars
```


```{r}
mtcars_corr_matrix <- cor(mtcars[-1])
mtcars_corr_matrix
```

```{r}
mtcars_corr_matrix_r_sorted <- sort(abs(mtcars_corr_matrix))
mtcars_corr_matrix_r_sorted
```

## Question 2

Create the linear model: wt + hp + wt*hp and look at the QQ-plot.  In this plot, Normality seems like a reasonable assumption?

```{r}
mtcars_mlr1 <- lm(mpg ~ wt + hp + wt*hp, data = mtcars)
plot(mtcars_mlr1)
```

## Question 3

Using the model from above, create the residual plot.  What would be your conclusion about the assumption of homogeneity of the variance?

Answer: Homogeneity of the variance appears appropriate since the spread of the residuals are fairly consistent.

## Question 4

We can use the residual plot to help us decide if the errors are Normally distributed.

Answer: False

# Question Set 2

For the rest of the problems, we will be using the cars2010 data set (this data set is a training data set, so you do not need to subset....use the entire data set).

```{r}
cars <- read_csv("https://raw.githubusercontent.com/IAA-Faculty/statistical_foundations/refs/heads/master/cars2010.csv")
```

## Question 5

Create the linear model with the following variables and calculate the Cook's D statistic.  What is the largest value for Cook's D (recall that FE..Fuel Economy is the response variable)? Keep answer accuracy to three decimal places.

```{r}
cars_mlr <- lm(FE ~ EngDispl + C(Transmission) + C(AirAspirationMethod) + C(TransLockup) + C(TransCreeperGear) + C(DriveDesc) + C(IntakeValvePerCyl) + C(CarlineClassDesc) + C(VarValveLift), data = cars)

cars_cook <- cooks.distance(cars_mlr)
sort(cars_cook, decreasing = TRUE)
```

## Question 6

Now let's use the Hat values to identify influential points and external studentized residuals to identify outliers.  How many observations are above the Hat value cutoff AND has a magnitude above 3?

```{r}
c.index = seq(1, nrow(cars))

h_threshold <- 20/nrow(cars)
h_bool = hatvalues(cars_mlr)>h_threshold
h_plot = ggplot(cars_mlr, aes(x = c.index, y=hatvalues(cars_mlr))) + geom_point() + geom_line(y=h_threshold)
plot(h_plot)

rstudent_bool = abs(rstudent(cars_mlr))>3
rstudent_plot = ggplot(cars_mlr, aes(x = c.index, y=rstudent(cars_mlr))) + geom_point() + geom_line(y=3) + geom_line(y=-3)
plot(rstudent_plot)

both_bool = h_bool & rstudent_bool
num_influence = sum(both_bool)
num_influence
```

```{r}


```


























