---
title: "Stats Lab 12 - Categorical Associations"
output: html_notebook
---

```{r}
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(gmodels)
library(epitools)
library(DescTools)
library(vcd)
library(vcdExtra)
install.packages('vcdExtra')
```

# Question Set 1

For this question, we will be using the safety dataset. Data were collected in an effort to relate the safety of certain vehicles to different aspects of vehicles. 

This dataset has the following variables:
Unsafe: binary safety designations (1 = below average/unsafe, 0 = average or above average/safe)
Type: type of care (large, medium, small, sport/utlilty, sports)
Region: manufacturing region(Asia, N. America)
Weight: Integer value for car weight ranging from 1 to 6
Size: size of car corresponding to Tpye (1 = small/sports, 2 = medium, 3 = large or sport/utility)

Do not create a training and testset, just use the whole dataset for the following analysis.

Examine the association between Region and Unsafe.

```{r}
safety <- read_csv('https://raw.githubusercontent.com/IAA-Faculty/statistical_foundations/refs/heads/master/safety.csv')
spec(safety)
```

## Question 1a

What percentage of cars manufactured in Asia were classified as unsafe? Round to 1 decimal (keep in mind, the question asks for a percentage, not proportion)

```{r}
CrossTable(safety$Unsafe, safety$Region)
```

Answer: 57.1%

## Question 1b

What percentage of cars classified as safe were manufactured in North America?

Answer: 50.0%

## Question 1c

What are the odds that a car manufactured in Asia is safe (remember the definition of the safety variable... 1=unsafe)? Round to 2 decimals

```{r}
0.571/0.429 
```

## Question 1d

What are the odds that a car manufactured in North America is safe? Round to 3 decimals

```{r}
0.754/0.246
```

## Question 1e

What are the odds of a car manufactured in Asia being safe vs a car manufactured in North America? Round to 4 decimals

```{r}
asia_safe <- 0.571/0.429
na_safe <- 0.754/0.246
asia_safe / na_safe
```

# Question Set 2

We will be using the bike dataset for this question. Use the following code to obtain the train/test dataset split:

set.seed(123)
bike <- bike %>% mutate(id = row_number())
train <- bike %>% sample_frac(0.7)
test <- anti_join(bike, train, by = 'id')

We will be using the following variables:

There are abnormal times where the number of casual users is greater than or equal to the number of registered users. You can use the following code to create a variable casual_high that captures this:

train$casual_high <- train$casual >= train$registered

```{r}
bike <- read_csv('https://raw.githubusercontent.com/IAA-Faculty/statistical_foundations/refs/heads/master/bike.csv')
```

```{r}
set.seed(123)
bike <- bike %>% mutate(id = row_number())
train <- bike %>% sample_frac(0.7)
test <- anti_join(bike, train, by = 'id')
```

```{r}
train$casual_high <- train$casual >= train$registered
```

## Question 2a

You want to know if the occurrence of these times is related to the season of the year. Even though season is ordinal, you want to just test a general association and not a linear one, so you use a Pearson Chi-square test. What is the Chi-square value? round to 2 decimals

```{r}
assocstats(table(train$casual_high, train$season))
```

## Question 2b

You want to know if the occurrence of these times is related to the season of the year. Even though season is ordinal, you want to just test a general association and not a linear one, so you use a Pearson Chi-square test. What is the Chi-square value? round to 2 decimals

Answer: Reject the null hypothesis

## Question 2c

If you were to perform a Mantel-Hanszel Chi-Square test, what is the chi-square value?

```{r}
CMHtest(table(train$casual_high, train$season))$table[1,]
```

## Question 2d 

What is the result at significance level .001?

Answer: Fail to reject the null hypothesis

## Question 2e

You also want to know if the occurrence of these times is related to whether or not the day is a holiday. Perform the appropriate chi-square test to test this association. What is the result at significance level .001?

```{r}
assocstats(table(train$casual_high, train$holiday))
```








