---
title: "R Lab 2 - Data Structures"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

## Question 1

 Make a vector, "V1", of 100 elements starting from 10 and incremented by 10. What is the total value (sum) of "V1"?

```{r}
V1 <- seq(from = 10, by = 10, length.out = 100)
sum(V1)
```
## Question 2

Make a vector, "V2", of 100 elements of the value 25. What is the total value (sum) of "V2"?

```{r}
V2 <- rep(25, 100)
sum(V2)
```
## Question 3

Make a vector, "V3", that is "V1" divided by "V2". What is the max value of "V3"?

```{r}
V3 <- V1/V2
max(V3)
```
## Question 4

 Create data frame, "df1", comprised of "V1","V2", and "V3". Name the columns "Vector_1", "Vector_2", and "Vector_3" respectively. Paste your code in the text box below.

```{r}
df1 <- data.frame(
  "Vector_1" = V1,
  "Vector_2" = V2,
  "Vector_3" = V3
)
head(df1)
```
## Question 5

How many rows are in "df1"?

```{r}
dim(df1)
```
## Question 6

What is the index of the max observation in "df1" column "Vector_3"?

```{r}
which.max(df1$Vector_3)

```
## Question 7

Create a new column in "df1", "Vector_4", that is the result of adding "Vector_1" and "Vector_3" together. What is the maximum value in "Vector_4"?

```{r}
df1$Vector_4 <- df1$Vector_1+df1$Vector_2
max(df1$Vector_4)
```
## Question 8

Move the column "Vector_4" to the first column of "df1" Paste your code in the text box below.

```{r}
df1 <- df1[,c(4,1:3)]
```

## Question 9

What value is in row 81, column 1 of "df1"?

```{r}
df1[81,1]
```
## Question 10

Reassign the minimum value in row 86 to 1. What is the mean of "Vector_2"?

```{r}
df1[86,which.min(df1[86,])] <- 1
mean(df1$Vector_2)
```
## Question 11

Create a new column in "df1", "Row_Mean", that is the average of the values in columns. What is the value of "Row_Mean" for row 45?

```{r}
df1$Row_Mean <- rowMeans(df1[c(1:)])
df1[45,"Row_Mean"]
```
## Question 12

Create a list, "l1", from "df1". Paste your code in the text bow below.

```{r}
l1 <- as.list(df1)
l1
```
## Question 13

What is the mean of the combined elements from "l1" components "Vector 1" and "Vector 2"?

```{r}
mean(c(l1$Vector_1,l1$Vector_2))
```