---
title: "R Programming Lab 1"
output: html_notebook
---

## Question 1

What does the log() function do?

```{r}
log(20)
```
Answer: It returns the value of the natural log of the value called with the function

## Question 2

Is the default calculation for the log function natural log or base 10?

```{r}
log(10)
```
Answer: Natural Log

## Question 3

What are the arguments for the log() function?

```{r}
?log
```

Answer: log() has two arguments: x, a numeric or complex vector, and base, a positive or complex number: the base with respect to which logarithms are computed. Defaults to e=exp(1).

## Question 4

What is the log of 1.4, base 8.3 (all digits)?

```{r}
log(x=1.4, base=8.3)
```

## Question 5

Assign the result of question 4 to the variable "result". Copy code statement in answer box below.

```{r}
result <- log(x=1.4, base=8.3)
```

## Question 6

Assign the variable "result" to "result" rounded to thousandth. What is the value of result?

```{r}
result <- round(result, 3)
result
```

## Question 7

What is the value of "result" mod 0.05?

```{r}
result %% 0.05
```

## Question 8 

Assign the value 5 to the variable "value". Copy the code statement in the answer box below.

```{r}
value <- 5
```

## Question 9

What is "value" divided by "result"?  Please include 5 decimals in your solution

```{r}
round(value/result,5)
```

## Question 10

What is the ceiling of "value" divided by "result", divided by the floor of "value" divided by "result" (all digits)?

Follow the order of operations:
- ceiling of value divided by result
- divided by
- floor of value divided by result

```{r}
num <- ceiling(value/result)
denom <- floor(value/result)
ans <- num/denom
ans
```

## Question 11

Attach or copy your code below
