---
title: "Lab 4 - Data Wrangling"
output: html_notebook
---

```{r}
install.packages(c("data.table","dplyr","stringr"))
library(data.table)
library(dplyr)
library(stringr)
```

## Question 1

Load the Auto_MPG.csv into a data frame named "df_auto". How many columns are in "df_auto"?

```{r}
df_auto <- read.csv("Auto_MPG.csv", stringsAsFactors = FALSE)
dim(df_auto)
```

## Question 2

Convert "Horsepower" to numeric. How many NAs are in "df_auto" after converting "Horsepower"?

```{r}
df_auto$Horsepower <- as.numeric(df_auto$Horsepower)
summary(df_auto)
df_auto$Horsepower
summary(df_auto$Horsepower)
```

## Question 3

Using dplyr, add to "df_auto":
- Variable named "Car_Make" that is the first word from "Car_Name"
- Variable named "Car_Model" that is the remaining words (not first) from "Car_Name"

How many observations of "Car_Make" "ford" are there?

```{r}
df_auto <- df_auto %>%
  rowwise() %>% 
  mutate(Car_Make=str_split(Car_Name, " ")[[1]][1]) %>%
  mutate(Car_Model=paste(str_split(Car_Name, " ")[[1]][-1], collapse=" "))

nrow(filter(df_auto, Car_Make == 'ford'))

```

## Question 4

Using dplyr, how many cars are there for model year 73?

```{r}
nrow(filter(df_auto, Model_Year == 73))
```

## Question 5

Using dplyr, what "Car_Make" "ford" car has the highest "Weight_per_Horsepower" (Weight/Horsepower) ratio? Ignore NAs.

```{r}
df_auto <- mutate(df_auto, Weight_per_Horsepower = Weight/Horsepower)

df_auto %>% filter(Car_Make == 'ford', na.rm = TRUE) %>% arrange(desc(Weight_per_Horsepower))
```

## Question 6

Using dplyr and stringr, update "df_auto" so all values of "Car_Make" that contain "chev" say "chevrolet".

How many observations have "Car_Make" "chevrolet"?

```{r}
df_auto <- mutate(df_auto, Car_Make = ifelse(str_detect(Car_Make,'chev'), 'chevrolet', Car_Make))

nrow(filter(df_auto, Car_Make == 'chevrolet'))

```

## Question 7

What is the average MPG for "chevrolet" cars?

```{r}
df_auto_chevy <- filter(df_auto, Car_Make == 'chevrolet')

mean(df_auto_chevy$MPG)

```

## Question 8

Below is a data frame, "df_auto_start", that contains the first production year for some of the auto makes.

  Car_Make <- c("amc", "audi", "bmw", "buick", "chevrolet", "datsun", "dodge", "ford")
  First_Year <- c(1954, 1910, 1916, 1903, 1911, 1931, 1900, 1903)
  df_auto_start <- data.frame(Car_Make=Car_Make, First_Year=First_Year, stringsAsFactors = FALSE)

Using dplyr joins, how many cars exist in the data set where "First_Year" is null?

```{r}

```



