---
title: "R Lab 3 - Programming Fundamentals"
output: html_notebook
---

## Question 1

Set the seed to 1. Paste the correct code statement in the box below.

```{r}
set.seed(1)
```

## Question 2

Create a variable "pop_truth" that is 100,000 samples from an exponential distribution with rate 2.

What function is used to draw a random sample from an exponential distribution?

```{r}
pop_truth <- rexp(100000, rate = 2)
```

## Question 3

What is the mean of "pop_truth", rounded to two decimals?

```{r}
mean(pop_truth)
```

## Question 4

Create a variable "sample_pop" that is a random sample of 5,000 values from "pop_truth". Before running your code, set the seed to 1 again. Note, the seed needs to be set each time you call a function with a random component. Paste your code in the text box below.

```{r}
set.seed(1)
sample_pop <- sample(pop_truth, 5000)
```

## Question 5

Create a function named "sample_stats" that calculates and returns the mean and standard deviation from a provided vector. Paste your code in the text box below.

```{r}
sample_stats <- function(vec){
  ans <- list(mean=mean(vec), std=sd(vec))
  return(ans)
}
```

## Question 6

Use the function to determine the mean and standard deviation from "sample pop". What is the returned mean, rounded to two decimals?

```{r}
sample_stats(sample_pop)
```

## Question 7

7 Create a for loop from 1 to 10,000 (note, use a smaller value when developing) that does the following:

\# 7.1 - If the first iteration, create a new data frame named "df_results" with three columns - "mean", "stdev", and "iteration"

\# 7.2 - Draws a new 500 observation sample from "pop_truth" to "sample_pop"

\# 7.3 - Uses the "sample_stats" function created previously to calculate and return the mean and standard deviation of "sample_pop" to a new variable named "metrics"

\# 7.4 - Adds a third element to "metrics" that is the current value of the iterator

\# 7.5 - Appends "metrics" to the data frame "df_results"

\# 7.6 - Prints the message, "Currently on iteration" and the current value of the iterator

Paste your code to the text box below.

```{r}
set.seed(1)

df_results <- data.frame(
  mean = c(), 
  stdev = c(),
  iteration = c()
)

for(i in 1:20){
  sample_pop <- sample(pop_truth, 500)
  metrics <- sample_stats(sample_pop)
  
  metrics_df = data.frame(mean = metrics$mean,
                          stdev = metrics$std,
                          iteration = i)
  
  df_results <- rbind(df_results, metrics_df)
  print(paste("Currently on iteration ", i))
}
```

## Question 8

What is the average of "df_results\$mean", rounded to two decimals?

```{r}
mean(df_results$mean)
```

## Question 9

Compute the 5th and 95th percentiles of "df_results\$mean". What is the 5th percentile, rounded to two decimals?

```{r}
quantile(df_results$mean, probs = c(0.05, 0.95))
```

## Question 10

Use the hist() function to create a histogram of "df_results\$mean" and "pop_truth". What type of distribution does "df_results\$mean" appear to be?

```{r}
hist(df_results$mean)
```

```{r}
hist(pop_truth)
```
